\chapter{Project Planning}

\textit{This section contains planning constraints, a description of the different project phases and the actual project planning.}

\section{Planning Constraints}
The following constraits apply to the project planning.

\subsection{Time Constraint}
The project consists of two parts, namely the Research Topics part and the graduation project part. The Research Topics part takes about 10EC and the graduation project part takes 30EC in total. Every European Credit (EC) stands for 28 hours, so the total time for Research Topics is $10*28 = 280$ hours and for the graduation project $30 * 28 = 840$ hours.

The graduation project has to be completed strictly within half a year, which is 26 weeks. The research topics does not have a fixed time constraint. This time constrait can be elevated a bit, but this requires permission from the project committee. 

\subsection{Unavailability due to Work}
During the research project, Wytse is available for four days a week. This means that Wytse is unavailable one day a week and this extends the length of the graduation project to $1.2 * 26 \approx 32$ weeks.

\subsection{Research Honours Programme}
Wytse is planning to participate in the Research Honours program, which takes an extra $20EC$ if he succeeds. This causes the graduation project to take $50EC$ in total.

\subsection{Winter and Summer Holidays}
The winter holidays, which are two weeks, are excluded from the planning. In case the project takes $50EC$ or more, it will span the summer holidays. In that case the summer holidays are also excluded from the planning.

\subsection{Starting date}
We make use of the fact that the Research Topics phase does not have a fixed deadline. The prerequisite for starting the project is that software for PGAS/MPI is installed on the cluster, so we do not start with the project until the software is installed. We plan to start the project on November 3, 2014.

\subsection{Date of Completion}
Unfinished..

\section{Project Phases}
Below are the different project phases and milestones planned throughout the graduation project. Also a short indication is given of the time it takes to complete the phase.

\subsection{Microbenchmarking RDMA}
One of points mentioned in the project risks section of this report is the possible performance drop when increasing the amount of RDMA-exposed memory. This needs to be tested as early as possible, so the project starts by microbenchmarking the performance of RDMA.  

\subsection{Implementing and Microbenchmarking the Distributed Hash Table}
Developing the distributed hash table on top of MPI/PGAS takes about a month to complete. Microbenchmarking the implementation and discussing the results takes about two weeks. If the experimental results disappoint, some extra time is needed to deal with problems in the implementation.

\subsection{Implementing and Microbenchmarking the Work Stealing Framework}
Implementing and testing work stealing also takes a month in total. The scalability of the work stealing framework has to be tested, since it gives an indication of the scalability of the BDD operations. This can be done by performing a microbenchmark. If the benchmark results dissapoint, another work stealing framework can be tried or changes can be applied to the framework.

\subsection{Implementing and Benchmarking the BDD operations}
After having a good implementation of the distributed hash table and a work stealing protocol, the BDD operations have to be implemented. This will take roughly three weeks and benchmarking the BDD operations takes about two weeks.

\subsection{Finishing the Thesis}
Some time is reserved just for finishing the master thesis. Note that part of the writing process can be done during the other phases. We are planning to also release a paper on the results of this research project, so also the paper is finished during this phase.

\section{Planning}
The project planning is presented in Table \ref{tab:planning1} and \ref{tab:planning2}.

\begin{table}[ht]
	\begin{adjustwidth}{-.5in}{-.5in} 
	\centering
	\begin{tabular}{| c | c | c | l |}
		\hline
		\textbf{\specialcell{Project\\week}} & \textbf{\specialcell{Date\\(begin - end)}} & \textbf{\specialcell{Nr. of\\hours}} & \textbf{Project activities} \\ 
		\hline \hline
		1 & \specialcell{Nov 3, 2014\\Nov 7, 2014} & 32 & Implementing the DHT, \textit{32 hours} \\ \hline
		2 & \specialcell{Nov 10, 2014\\Nov 14, 2014} & 32 & Implementing the DHT, \textit{32 hours} \\ \hline
		3 & \specialcell{Nov 17, 2014\\Nov 21, 2014} & 32 & \specialcell{Implementing the DHT, \textit{16 hours}\\Writing a chapter on the DHT, \textit{16 hours}} \\ \hline
		4 & \specialcell{Nov 24, 2014\\Nov 28, 2014} & 32 & \specialcell{Finishing the DHT, \textit{8 hours}\\Finishing the chapter on the DHT, \textit{24 hours}} \\ \hline \hline

		5 & \specialcell{Dec 1, 2014\\Dec 5, 2014} & 32 & Microbenchmarking the DHT, \textit{32 hours} \\ \hline
		6 & \specialcell{Dec 8, 2014\\Dec 12, 2014} & 32 & Microbenchmarking the DHT, \textit{32 hours} \\ \hline
		7 & \specialcell{Dec 15, 2014\\Dec 19, 2014} & 32 & \specialcell{Adjusting the DHT, \textit{16 hours}\\Writing a chapter with benchmark conclusions, \textit{16 hours}} \\ \hline
		8 & \specialcell{Jan 5, 2015\\Jan 9, 2015} & 32 & \specialcell{Adjusting the DHT, \textit{8 hours}\\Finishing the chapter on benchmark conclusions, \textit{24 hours}} \\ \hline \hline

		9 & \specialcell{Jan 12, 2015\\Jan 16, 2015} & 32 & Implementing work-stealing, \textit{32 hours} \\ \hline
		10 & \specialcell{Jan 19, 2015\\Jan 23, 2015} & 32 & \specialcell{Implementing work-stealing, \textit{16 hours}\\Writing a chapter on work-stealing, \textit{16 hours}} \\ \hline
		11 & \specialcell{Jan 26, 2015\\Jan 30, 2015} & 32 & \specialcell{Testing work-stealing, \textit{16 hours}\\Finishing the chapter on work-stealing, \textit{16 hours}} \\ \hline \hline

		12 & \specialcell{Feb 2, 2015\\Feb 6, 2015} & 32 & Implementing BDD operations, \textit{32 hours} \\ \hline
		13 & \specialcell{Feb 9, 2015\\Feb 13, 2015} & 32 & Implementing BDD operations, \textit{32 hours} \\ \hline
		14 & \specialcell{Feb 16, 2015\\Feb 20, 2015} & 32 & Implementing BDD operations, \textit{32 hours} \\ \hline
		15 & \specialcell{Feb 23, 2015\\Feb 27, 2015} &32 & \specialcell{Implementing BDD operations, \textit{16 hours}\\Writing a chapter on BDD operations, \textit{16 hours}} \\ \hline
		16 & \specialcell{Mar 2, 2015\\Mar 6, 2015} & 32 & \specialcell{Implementing BDD operations, \textit{16 hours}\\Writing a chapter on BDD operations, \textit{16 hours}} \\ \hline
		17 & \specialcell{Mar 9, 2015\\Mar 13, 2015} & 32 & \specialcell{Finishing the BDD operations, \textit{8 hours}\\Writing a chapter on BDD operations, \textit{24 hours}} \\ \hline
	\end{tabular}
	\end{adjustwidth}
	\caption{Graduation project planning, first part, weeks 1 to 17.}
	\label{tab:planning1}
\end{table}

\begin{table}[ht]
	\begin{adjustwidth}{-.5in}{-.5in} 
	\centering
	\begin{tabular}{| c | c | c | l |}
		\hline
		\textbf{\specialcell{Project\\week}} & \textbf{\specialcell{Date\\(begin - end)}} & \textbf{\specialcell{Nr. of\\hours}} & \textbf{Activities} \\ 
		\hline \hline
		18 & \specialcell{Mar 16, 2015\\Mar 20, 2015} & 32 & Benchmarking the BDD operations, \textit{32 hours} \\ \hline
		19 & \specialcell{Mar 23, 2015\\Mar 27, 2015}  & 32 & \specialcell{Benchmarking the BDD operations, \textit{24 hours}\\Writing a chapter on benchmark results, \textit{8 hours}} \\ \hline
		20 & \specialcell{Mar 20, 2015\\Apr 3, 2015}  & 32 & \specialcell{Benchmarking the BDD operations, \textit{16 hours}\\Writing a chapter on benchmark results, \textit{16 hours}} \\ \hline
		21 & \specialcell{Apr 6, 2015\\Apr 10, 2015}  & 32 & \specialcell{Finishing benchmarking, \textit{8 hours}\\Finishing the chapter on benchmark results, \textit{24 hours}} \\ \hline \hline

		22 & \specialcell{Apr 13, 2015\\Apr 17, 2015} & 32 & \specialcell{Writing a concept thesis, \textit{32 hours}} \\ \hline
		23 & \specialcell{Apr 20, 2015\\Apr 24, 2015} & 32 & \specialcell{Writing a concept thesis, \textit{32 hours}} \\ \hline
		24 & \specialcell{Apr 27, 2015\\May 01, 2015} & 32 & \specialcell{Writing a concept thesis, \textit{8 hours}\\Writing the paper, \textit{24 hours}} \\ \hline
		25 & \specialcell{May 04, 2015\\May 08, 2015} & 32 & \specialcell{Writing a concept thesis, \textit{8 hours}\\Writing the paper, \textit{24 hours}} \\ \hline
		26 & \specialcell{May 11, 2015\\May 15, 2015} & 32 & \specialcell{Writing a concept thesis, \textit{16 hours}\\Writing the paper, \textit{16 hours}\\Handing in the concept thesis} \\ \hline \hline

		27 & \specialcell{May 18, 2015\\May 22, 2015} & 32 & \specialcell{Creating a presentation, \textit{32 hours}} \\ \hline
		28 & \specialcell{May 25, 2015\\May 29, 2015} & 32 & \specialcell{Preparing the presentation, \textit{8 hours}\\ Lunch-meeting presentation, \textit{2 hours}\\Finishing the thesis, \textit{20 hours}} \\ \hline \hline

		29 & \specialcell{Jun 01, 2015\\Jun 05, 2015} & 32 & \specialcell{Finishing the thesis, \textit{32 hours}} \\ \hline
		30 & \specialcell{Jun 08, 2015\\Jun 12, 2015} & 32 & \specialcell{Finishing the thesis, \textit{8 hours}\\Finishing the paper, \textit{24 hours}} \\ \hline

		31 & \specialcell{Jun 15, 2015\\Jun 20, 2015} & 32 & \specialcell{Finishing the thesis, \textit{16 hours}\\Adjusting the presentation, \textit{16 hours}} \\ \hline
		32 & \specialcell{Jun 22, 2015\\Jun 26, 2015} & 32 & \specialcell{Handing in Thesis\\Preparing the presentation, \textit{24 hours}\\Presenting the thesis, \textit{8 hours}} \\
		\hline
	\end{tabular}
	\end{adjustwidth}
	\caption{Graduation project planning, second part, weeks 18 to 32.}
	\label{tab:planning2}
\end{table}

\subsection{Planning Risks}
The most important aspect of this project is the creation and evaluation of a distributed hash table that uses PGAS and one-sided RDMA operations where possible. If the implementation of the distributed hash table takes longer than expected for some reason, the implementation of BDD operations can be omitted. If needed, the paper can also be written after the graduation project.

The implementation of the work stealing protocol can be done ourselves, but if time runs short Scioto \cite{dinan2008scioto} can be used instead. Scioto is a work-stealing framework especially designed for the PGAS model and exploits locality. 

